A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning

### 2016-07-15 paper.tex
phrasing

  \paragraph{The free-energy / utility principle (of Friston ?).}
  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$. We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.  The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$. The difference in free-utility is then
  \begin{equation}
    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity of change}},
    \label{eq:free}
  \end{equation}
  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
  In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
a trade-off between an expected utility $\Delta U$ (first term) and
the information cost of transforming Pi
into $P_{\text{fin}}$ (second
term). In this interpretation $P_{\text{ini}}$ represents an initial choice
probability or policy, which includes the special case of the
uniform distribution where the decision-maker has initially no
preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
considers the utility constraint U∗
that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
\begin{equation}
P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
  \end{equation}

>>>>

  \paragraph{The free-energy / utility principle (of Friston ?).}
  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$. We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.  The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$. The difference in free-utility is then
  \begin{equation}
    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
    \label{eq:free}
  \end{equation}
  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
  In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
a trade-off between an expected utility $\Delta U$ (first term) and
the information cost of transforming Pi
into $P_{\text{fin}}$ (second
term). In this interpretation $P_{\text{ini}}$ represents an initial choice
probability or policy, which includes the special case of the
uniform distribution where the decision-maker has initially no
preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
considers the utility constraint U∗
that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
\begin{equation}
P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
  \end{equation}
